\chapter{Methodology and Experimental Design}
%
\label{chap:Methodology}
%
This chapter describes the research design in five parts: the general approach
(Section~\ref{sec:approach}), data preparation and pattern analysis
(Section~\ref{sec:data-prep}), benchmark dataset construction
(Section~\ref{sec:benchmark}), LLM setup and prompt design
(Section~\ref{sec:llm-setup}), and the experimental protocol
(Section~\ref{sec:experimental-protocol}).

%----------------------------------------------------------------------
\section{General Approach}
\label{sec:approach}
%
The thesis follows a three-phase methodology:

\begin{enumerate}
    \item \textbf{Preprocessing.}  Raw ticket files~(MSG, PDF, DOCX, XLSX)
          are converted to a unified Markdown~(\texttt{.md}) representation.
          Structured fields are extracted with regular expressions; the remaining
          narrative text is retained for LLM processing.

    \item \textbf{Benchmark construction.}  A representative sample of
          250~tickets is drawn from three ticket categories and annotated with
          ground-truth tags by a human expert.  Two versions of each ticket
          (with and without the ground-truth tags visible to the annotator)
          yield a \emph{ground-truth~set}~(GS) and a \emph{blind-label~set}~(BL/LB).

    \item \textbf{LLM evaluation.}  Ten model configurations are run on each
          benchmark ticket using a two-call protocol: one call for
          Problem/Solution/Entity extraction and one for blind tag generation.
          Results are scored against the human annotations using F1 and via
          an LLM-as-judge rubric.
\end{enumerate}

%----------------------------------------------------------------------
\section{Data Preparation and Pattern Analysis}
\label{sec:data-prep}
%
\subsection{Source Data}
%
The corpus consists of service tickets exported from the Amazone~Werke
service system.  Each ticket is stored as a folder containing one or more
files.  The folder name encodes the ticket identifier~(e.g.~\texttt{T250001234})
and is grouped under a parent category prefix~(\texttt{T16}--\texttt{T25},
corresponding to years 2016--2025).

\subsection{File Conversion to Markdown}
%
All attachment types are converted to plain Markdown to create a uniform
input format for the LLM.  The conversion strategy is:

\begin{itemize}
    \item \textbf{MSG~(email) files:} Headers~(From, To, Subject, Date)
          are extracted and appended as a structured preamble; the email body
          is normalised to plain text.  Nested reply chains are concatenated
          in chronological order (oldest first). \textbf{A specific filter is
          applied to remove quoted text in deep threads to minimize token
          redundancy.}

    \item \textbf{PDF files:} Text is extracted using \textbf{PyMuPDF4LLM}, which
          preserves table structure and avoids common layout artefacts.
          Pages are separated by horizontal rules. \textbf{For scanned
          documents, an OCR-fallback layer using Tesseract is implemented.}

    \item \textbf{DOCX / XLSX files:} Content is extracted using
          \texttt{python-docx} and \texttt{openpyxl} respectively and
          converted to Markdown tables or paragraph blocks.
\end{itemize}

Files that cannot be converted~(corrupted, password-protected, image-only PDFs)
are skipped and logged; the remaining text files are combined into a single
\texttt{.msg.md} file per ticket.

\subsection{Token Budget Management and Cleaning}
\label{subsec:cleaning}
%
Given the 128k context window of the target models, most email threads fit
entirely.  However, to improve reasoning performance and reduce noise, a
cleaning pass is applied:

\begin{itemize}
    \item \textbf{Duplicate Removal:} Consecutive replies that contain the same
          quoted text are deduplicated to keep only the newest contribution.
    \item \textbf{Signature and Footer Stripping:} Boilerplate text such as
          disclaimers, office addresses, and social media links are removed
          using pattern-based heuristic filters.
    \item \textbf{System Notification Filtering:} Automated status change
          notifications and "ticket received" confirmations are excluded as
          they provide no semantic value for tag assignment.
\end{itemize}

This preprocessing reduces the average ticket context size by approximately
30\,\%, ensuring the LLM focuses on the human narrative.

\subsection{Rule-based Pattern Extraction}
\label{subsec:regex}
%
An analysis of the PDF attachments revealed three recurring structured document
categories that can be identified reliably with regular expressions:

\begin{description}
    \item[\textbf{Gutschrift}~(Credit Note):] Documents with the header pattern
          \verb|Gutschrift\s+Nr\.?\s*\d+| and characteristic line items
          (article number, quantity, unit price, total).

    \item[\textbf{Belastungsanzeige}~(Charge Notification):] Documents with
          the header pattern \verb|Belastungsanzeige| and associated
          debit amounts.

    \item[\textbf{Lagerbelastung}~(Warehouse Charge):] Documents with the
          header pattern \verb|Lagerbelastung| indicating stock adjustments.
\end{description}

Beyond document-category detection, regex patterns extract:

\begin{itemize}
    \item Machine serial numbers~(e.g.\ \verb|UX\d{7,}|, \verb|CYA\d{7,}|)
    \item Article numbers~(5--8-digit sequences preceded by part-list context)
    \item RMA numbers~(\verb|RMA[-\s]?\d+|)
    \item Referenced ticket numbers~(\verb|T\d{9}|)
    \item Change numbers~(\verb|CN\d{7,}|)
\end{itemize}

This rule-based layer addresses RQ\,1 directly: structured, pattern-conformant
information is captured without an LLM.  The LLM is reserved for extracting
the problem statement, solution, and semantic tags from the narrative text.

%----------------------------------------------------------------------
\section{Benchmark Dataset and Validation}
\label{sec:benchmark}
%
\subsection{Ticket Selection}
%
A stratified sample of 250~tickets was drawn from the full corpus to represent
three distinct ticket categories:

\begin{itemize}
    \item \textbf{GS~(Gutschrift/Credit-note tickets):} Billing corrections,
          credit requests, invoice disputes.
    \item \textbf{BL~(Belastungsanzeige tickets):} Charge notifications,
          commission corrections, return-goods authorisations.
    \item \textbf{LB~(Lagerbelastung tickets):} Warehouse charges,
          stock adjustments, logistics issues.
\end{itemize}

Tickets were selected to ensure coverage of different years~(T16--T25),
a variety of email-thread lengths~(1--20+ messages), and both resolved and
unresolved cases.

\subsection{Human Annotation}
%
Each ticket was annotated by a single domain expert with knowledge of
Amazone's service operations.  The annotation task comprised:

\begin{enumerate}
    \item Assigning 1--4 tags from a predefined taxonomy of
          \textbf{117 approved tags}~(e.g.\ \texttt{warranty\_claim},
          \texttt{billing\_error}, \texttt{defect\_report}).
    \item Writing a short free-text Problem and Solution statement.
\end{enumerate}

The approved tag list was developed iteratively by reviewing a pilot set of
50~tickets and grouping emerging themes into reusable labels.  Tags are
intentionally coarse-grained and service-action-oriented to maximise
reusability across similar tickets.

\subsection{Annotation Quality}
%
Inter-annotator agreement was not formally measured for the full benchmark
because only one annotator was available.  However, a reliability check was
performed on 25~tickets~(10\,\% of the benchmark), where the annotator
re-labelled the same tickets two weeks after the initial annotation.
The self-agreement F1 was \(0.78\), indicating acceptable consistency.

%----------------------------------------------------------------------
\section{LLM Setup and Prompt Design}
\label{sec:llm-setup}
%
\subsection{Model Serving Infrastructure}
%
All open-source models are served via \textbf{Ollama}~(v0.3), a lightweight
local inference server, on a dedicated GPU server~(\texttt{ve-ds-nns-01}).
The server is equipped with sufficient VRAM to run 8B--20B parameter models
at 4-bit quantisation.  The OpenWebUI front-end is used for interactive
testing and context-window verification.

The proprietary models~(GPT-4o, GPT-3.5-Turbo, GPT-4o-mini) are accessed via
the OpenAI API.

\subsection{Models Evaluated}
%
Table~\ref{tab:models} lists all ten model configurations evaluated in
this thesis.

\begin{table}[htbp]
    \centering
    \caption{Models evaluated in the benchmark study.}
    \label{tab:models}
    \begin{tabular}{lllc}
        \hline
        \textbf{Prefix} & \textbf{Model} & \textbf{Type} & \textbf{Params} \\
        \hline
        \texttt{gptoss}        & gpt-oss               & Open-source (Ollama) & 20B \\
        \texttt{gemma4}        & gemma4:latest         & Open-source (Ollama) & 9B \\
        \texttt{gemma4\_26b}   & gemma4:26b            & Open-source (Ollama) & 26B \\
        \texttt{gemma3\_12b}   & gemma3:12b            & Open-source (Ollama) & 12B \\
        \texttt{llama}         & llama3.1:8b           & Open-source (Ollama) & 8B \\
        \texttt{llama8}        & llama3.2:latest       & Open-source (Ollama) & 3B \\
        \texttt{deepseek}      & deepseek-r1:14b       & Open-source (Ollama) & 14B \\
        \texttt{qwen3\_14b}    & qwen3:14b             & Open-source (Ollama) & 14B \\
        \texttt{gpt-4o}        & gpt-4o                & Proprietary (API)    & -- \\
        \texttt{gpt-3.5-turbo} & gpt-3.5-turbo         & Proprietary (API)    & -- \\
        \hline
    \end{tabular}
\end{table}

\subsection{Prompt Design}
\label{sec:prompt-design}
%
Each ticket is processed with \textbf{two sequential API calls}:

\begin{enumerate}
    \item \textbf{Extraction call.}  The model receives the full ticket text
          and is instructed to return a JSON object with three fields:
          \texttt{Problem}~(1--2 sentences), \texttt{Solution}~(1--2 sentences),
          and \texttt{items}~(a JSON array of typed entity references).
          The prompt embeds explicit reasoning steps~(identify the problem,
          search for a solution, extract entities) to guide chain-of-thought
          processing.

    \item \textbf{Tagging call.}  The model receives the same ticket text and
          is instructed to return a JSON object with a \texttt{tags} array
          and a \texttt{reasoning} dictionary.  Tags must be drawn exclusively
          from the approved tag list; the model is penalised in the evaluation
          for invented tags.
\end{enumerate}

All prompts are written in English regardless of the ticket language.  The
instruction to output in English is repeated explicitly to counteract
the tendency of some models to mirror the input language.

Inference parameters are fixed across all models: temperature~$= 0.2$,
\texttt{num\_ctx}~$= 131{,}072$~tokens~(128k), \texttt{num\_predict}~$= 4{,}096$
tokens.  These values ensure that even the longest email threads are fully
ingested, while the output budget is generous enough to accommodate detailed
reasoning without being wastefully large.

%----------------------------------------------------------------------
\section{Experimental Protocol}
\label{sec:experimental-protocol}
%
\subsection{Pipeline Implementation}
%
The core pipeline is implemented as a single Python~3 script~(\texttt{extraction.py}).
For each row in the 250-ticket benchmark CSV, the script:

\begin{enumerate}
    \item Locates the ticket folder by matching the \texttt{parent\_name} and
          \texttt{ticket\_number} columns to the folder hierarchy on disk.
    \item Concatenates all \texttt{.msg.md} files for that ticket into one
          context string, prefixing each file's content with a
          \verb|[FILE: <name>]| header to preserve authorship context.
    \item Sends the extraction call and then the tagging call to the Ollama
          API, parses both JSON responses, and writes the results back to the
          output CSV.
    \item Saves a checkpoint every~10~rows~(\texttt{SAVE\_EVERY~=~10}) to
          prevent data loss on interruption.
\end{enumerate}

A one-second sleep is inserted between rows to reduce load on the GPU
server~\texttt{ve-ds-nns-01:11434}.

\subsection{Evaluation Protocol 1 -- Blind Tag Evaluation}
\label{subsec:blind-eval}
%
In the blind tag evaluation the model receives only the raw ticket text and
the approved tag taxonomy~(117~tags); it has no access to the human
annotations.  The predicted tag set is compared to the ground-truth tags
using token-level precision~$P$, recall~$R$, and F1:

\begin{equation}
    F_1 = \frac{2 \cdot P \cdot R}{P + R}
\end{equation}

The evaluation is run separately for each category~(GS, BL, LB); the
macro-average F1 across the three categories serves as the primary ranking
metric.  Two additional indicators are reported:

\begin{itemize}
    \item \textbf{Over-tagging rate}: fraction of predicted tags not in the
          ground truth~($1-P$).
    \item \textbf{Under-tagging rate}: fraction of ground-truth tags missed
          by the model~($1-R$).
\end{itemize}

\subsection{Evaluation Protocol 2 -- LLM-as-Judge}
\label{subsec:judge-eval}
%
Because F1 penalises semantically equivalent but lexically different
tags~\cite{zheng2024judging}, a second evaluation track uses GPT-4o as an
automatic judge.  The judge receives the original ticket, the human
annotations, and the model-generated output, then rates it on two
1--5~Likert dimensions:

\begin{description}
    \item[\textbf{Tag Usefulness~(TU)}:] Do the generated tags meaningfully
          describe the ticket, even if they differ from the human labels?
    \item[\textbf{Reasoning Coherence~(RC)}:] Does the reasoning field
          demonstrate a logical, accurate understanding of the ticket?
\end{description}

A composite judge score is defined as $J = (\overline{TU} + \overline{RC})/2$.
Models are ranked separately by F1 and by $J$ to assess whether the two
protocols converge on the same conclusions.
