\chapter*{Declaration}
%
I declare that I have written this thesis independently. I have acknowledged
all passages taken verbatim or in spirit from published or unpublished works
of others or from the author himself\slash herself. All sources and aids that
I have used for the thesis are indicated. The thesis has not been submitted
to any other examination authority with the same content or in essential
parts.

\subsection*{Declaration of Generative AI and AI-Assisted Technologies
             in the Writing Process}
%
I confirm that I have disclosed the use of all generative AI tools employed in
the preparation of this thesis, naming each tool and its purpose in the table
below.  I used these tools only as an aid; the substantive content, analysis,
and conclusions of this thesis are my own.  All AI-generated content was
critically reviewed, verified, and edited by me, and I take full responsibility
for the content of this thesis, including any passages that were drafted or
revised with AI assistance.  I am aware that the undisclosed use of generative
AI tools may be treated as an attempt to deceive under the relevant examination
regulations.

\begin{table}[H]
    \centering
    \renewcommand{\arraystretch}{1.3}
    \begin{tabular}{p{0.26\textwidth} p{0.28\textwidth} p{0.34\textwidth}}
        \hline
        \textbf{Tool (Provider)} & \textbf{Version} & \textbf{Purpose} \\
        \hline
        ChatGPT (OpenAI) & GPT-4 class, 2025 &
        Language editing and grammar checking of author-written text \\

        Claude (Anthropic) & Claude 4 class, 2025 &
        LaTeX formatting support and proofreading of author-written text \\
        \hline
    \end{tabular}
\end{table}

\noindent
The use of AI tools was limited to the purposes listed above.  No part of the
research design, data analysis, results, or their interpretation was delegated
to a generative AI system.

\vspace{2cm}
%
\begin{tabular}{@{}l@{}}%
\rule{0.45\textwidth}{0.4pt}\\
Place, date%
\end{tabular}%
\hfill%
\begin{tabular}{@{}l@{}}%
\rule{0.45\textwidth}{0.4pt}\\
Signature%
\end{tabular}%