\chapter{Results and Observations}
%
\label{chap:Results}
%
This chapter presents the quantitative and qualitative results of the
evaluation protocols described in Chapter~\ref{chap:Methodology}.
Section~\ref{sec:res-f1} reports token-level F1 scores comparing each model's
tags against the human ground truth.
Section~\ref{sec:res-judge} presents the LLM-as-judge ratings.
Section~\ref{sec:res-semantic} analyses semantic similarity between model
outputs and the ticket content.
Section~\ref{sec:res-oss} compares the open-source models,
Section~\ref{sec:res-fullrun} describes the large-scale run on the full archive, and
Section~\ref{sec:observations} summarises cross-cutting observations.

%----------------------------------------------------------------------
\section{Tag Evaluation Against Human Ground Truth: F1 Scores}
\label{sec:res-f1}
%
\subsection{All Models vs Human Tags Baseline}
\label{subsec:f1-all}
%
This section reports how closely each model's predicted tags match the
28~human-defined ground-truth tags, measured with the token-level F1 score.
Table~\ref{tab:f1-results} summarises the F1, Precision, Recall,
and Accuracy scores for all fifteen model configurations evaluated against
the human-annotated ground-truth tags.  All scores are macro-averaged over
the 250~benchmark tickets.

\begin{table}[H]
    \centering
    \caption[Metric comparison: all models vs human tags baseline]{Metric comparison: all models vs human tags baseline (Tags).
             Best result in each column is \textbf{bold}.}
    \label{tab:f1-results}
    \begin{tabular}{lcccc}
        \hline
        \textbf{Model} & \textbf{F1 Score} & \textbf{Precision} & \textbf{Recall} & \textbf{Accuracy} \\
        \hline
        gpt-5\_4-mini       & \textbf{0.3943} & \textbf{0.4457} & 0.3911          & 0.9362 \\
        gpt-5.5             & 0.3822          & 0.4451          & 0.3819          & 0.9379 \\
        gpt-4o-mini         & 0.3746          & 0.3899          & \textbf{0.4114} & 0.9285 \\
        gptoss              & 0.3562          & 0.3759          & 0.4075          & 0.9218 \\
        gemma4              & 0.3122          & 0.2627          & 0.4951          & 0.8889 \\
        gpt-4o              & 0.3034          & 0.3290          & 0.3099          & 0.9302 \\
        qwen3\_14b          & 0.2803          & 0.3430          & 0.2539          & \textbf{0.9412} \\
        llama3.1            & 0.2712          & 0.2611          & 0.3249          & 0.9080 \\
        qwen2.5             & 0.2660          & 0.3240          & 0.2488          & 0.9378 \\
        gpt-3.5-turbo       & 0.2524          & 0.2746          & 0.2534          & 0.9105 \\
        dolphin             & 0.2181          & 0.2230          & 0.3259          & 0.8821 \\
        deepseek7b          & 0.2137          & 0.1985          & 0.2818          & 0.9089 \\
        gemma7b             & 0.1980          & 0.1444          & 0.3809          & 0.8246 \\
        llama3.2            & 0.1783          & 0.1523          & 0.2457          & 0.8881 \\
        deepseek14          & 0.1644          & 0.1815          & 0.1849          & 0.8731 \\
        \hline
    \end{tabular}
\end{table}

The results reveal several noteworthy patterns.  In what follows,
\emph{over-tagging} refers to a model predicting tags that are not in the human
ground truth~(which lowers precision), and \emph{under-tagging} refers to a
model missing ground-truth tags~(which lowers recall).  First, proprietary models
(GPT-5.4-mini, GPT-5.5, GPT-4o-mini) occupy the top three positions,
followed by the open-source GPTOSS model at rank~4 with an F1 of~0.3562.
Second, \texttt{gemma4} achieves the highest recall~(0.4951) among all
models but at the cost of low precision~(0.2627), indicating severe
over-tagging.  Third, \texttt{qwen3\_14b} achieves the highest
accuracy~(0.9412) but at the cost of low recall~(0.2539), suggesting a
conservative tagging strategy.

\subsection{Performance of Rule-based Extraction}
\label{subsec:regex-res}
%
Table~\ref{tab:regex-results} shows the performance of the regular expression
patterns described in Section~\ref{subsec:regex} for structured field
extraction.

\begin{table}[H]
    \centering
    \caption{Performance of regex patterns for structured extraction.}
    \label{tab:regex-results}
    \begin{tabular}{lccc}
        \hline
        \textbf{Field Type} & \textbf{Precision} & \textbf{Recall} & \textbf{F1} \\
        \hline
        Machine Serial Number & 0.992 & 0.975 & 0.983 \\
        Article Number        & 0.964 & 0.941 & 0.952 \\
        RMA Number            & 0.981 & 0.924 & 0.951 \\
        Document Category     & 0.978 & 0.962 & 0.970 \\
        \hline
    \end{tabular}
\end{table}

The near-perfect precision for structured fields confirms that LLM inference
is not required for these elements, allowing the LLM budget to be spent
entirely on narrative reasoning.

\subsection{LLM Tagging Profiles}
\label{subsec:tagging-profiles}
%
Table~\ref{tab:tagging-profiles} characterises each model's tagging behaviour
in terms of volume and diversity.  These profiles were produced in an
exploratory run in which the models were deliberately given no instruction
about \emph{how many} tags to generate and no hints about the desired
granularity.  The intention was to observe each model's unconstrained
behaviour: left to its own devices, how many tags does a model produce, and
how varied are they?  In the table, total tags refers to the number of tags
produced across all processed tickets, unique tags indicates how many distinct
tag strings the model generated, and average tags per ticket is the mean tag
count per ticket.

\begin{table}[H]
    \centering
    \caption[LLM tagging profiles]{LLM tagging profiles: volume and diversity of generated tags, produced without any instruction on tag count or granularity.}
    \label{tab:tagging-profiles}
    \begin{tabular}{lrrrr}
        \hline
        \textbf{Model} & \textbf{Total Tags} & \textbf{Unique Tags} & \textbf{Tickets} & \textbf{Avg Tags/Ticket} \\
        \hline
        gemma3 (12b)      & 1884 &  71 & 236 & 7.98 \\
        llama3.2 (3b)     & 1708 &  95 & 214 & 7.98 \\
        gemma (7b)        & 1768 &  65 & 222 & 7.96 \\
        gpt-oss (20b)     & 1670 &  95 & 210 & 7.95 \\
        llama3.1 (8b)     & 1532 &  46 & 192 & 7.98 \\
        deepseek-r1 (7b)  & 1610 & 182 & 202 & 7.97 \\
        deepseek-r1 (14b) & 1498 & 161 & 188 & 7.97 \\
        mistral (7b)      & 1654 & 148 & 208 & 7.95 \\
        qwen2.5 (7b)      & 1400 &  29 & 176 & 7.95 \\
        \hline
    \end{tabular}
\end{table}

Even without an explicit instruction, all models settled on roughly eight tags
per ticket, which suggests this is a natural granularity for the task.  The
number of \emph{unique} tags, however, varies dramatically: \texttt{qwen2.5}
reuses a small vocabulary of only~29 distinct tags, whereas
\texttt{deepseek-r1~(7b)} produces~182.  A high unique-tag count signals that
the model invents ad hoc labels rather than converging on a consistent
vocabulary, and it was precisely this uncontrolled proliferation of labels
that motivated the quality-control and consolidation process described in
Section~\ref{sec:disc-tags}.

%----------------------------------------------------------------------
\section{LLM-as-Judge Evaluation}
\label{sec:res-judge}
%
\subsection{Item Extraction Baseline}
\label{subsec:item-extraction}
%
As a complementary evaluation, GPT-5.5 was used as a judge to assess how
well each model's extracted items (machine numbers, article numbers, etc.)
match the ticket content.  The metric is \textit{Average Target Hit Rate}:
the fraction of ground-truth items the model correctly identifies.

\begin{table}[H]
    \centering
    \caption[Item extraction: average target hit rate]{Item extraction: average target hit rate (GPT-5.5 as baseline judge).}
    \label{tab:item-extraction}
    \begin{tabular}{clc}
        \hline
        \textbf{Rank} & \textbf{Model} & \textbf{Avg Target Hit Rate} \\
        \hline
        1 & GPT-5.4 MINI    & 0.7737 \\
        2 & GPT-4o MINI     & 0.6169 \\
        3 & GPTOSS          & 0.5110 \\
        4 & QWEN3 14B       & 0.4603 \\
        5 & ITEM EXTRACTION (human) & 0.4313 \\
        6 & LLAMA3.1        & 0.3990 \\
        7 & GEMMA4          & 0.3989 \\
        \hline
    \end{tabular}
\end{table}

Notably, three LLM models (GPT-5.4~MINI, GPT-4o~MINI, GPTOSS) outperform
the human annotator at item extraction.  This suggests that LLMs can surface
structured references (machine serials, article codes) that human annotators
overlook when manually scanning long email threads.

\subsection{Tag Usefulness and Reasoning Quality}
\label{subsec:judge-composite}
%
In a second judge evaluation, GPT-5.5 was given 28~base tags and asked to
rate the Problem/Solution quality of each model on a 0 to 10 scale and to assess
the usefulness of the model's tags as a percentage.  The four open-source
models were then ranked on each metric, and a \emph{composite} score was
derived from these ranks and rescaled so that the best-performing model
maps to~100 and the worst to the lowest value.  The composite is therefore a
\emph{relative} ranking score among the four candidates, not an absolute
measure of correctness: a composite of~100 means a model ranked first overall,
not that it achieved perfect output.

\begin{table}[H]
    \centering
    \caption[GPT-5.5 judge results for the four open-source candidate models]{GPT-5.5 judge results for the four open-source candidate models:
             Problem/Solution quality, tag usefulness, and the relative composite ranking
             score. 665~extended tags were evaluated. Per-metric ranks are
             shown in parentheses.}
    \label{tab:judge-results}
    \begin{tabular}{clccc}
        \hline
        \textbf{Rank} & \textbf{Model} & \textbf{Problem/Solution Quality} & \textbf{Usefulness} & \textbf{Composite (rel.)} \\
        \hline
        1st & GPTOSS      & 6.09/10 (R1) & 65.6\% (R1) & 100.0 \\
        2nd & GEMMA4       & 4.54/10 (R3) & 39.6\% (R2) &  62.5 \\
        3rd & QWEN3\_14B   & 5.27/10 (R2) & 29.2\% (R4) &  50.0 \\
        4th & LLAMA3.1     & 4.09/10 (R4) & 37.2\% (R3) &  37.5 \\
        \hline
    \end{tabular}
\end{table}

Among the four open-source models evaluated by the judge, GPTOSS ranks first on
both the Problem/Solution quality~(6.09/10) and the usefulness~(65.6\,\%)
metric, giving it the top relative composite score.  Its lead is driven mainly
by the large usefulness margin over the runner-up GEMMA4~(65.6\,\% versus
39.6\,\%).  Notably, QWEN3\_14B ranks second on Problem/Solution quality~(5.27) but only
fourth on usefulness~(29.2\,\%), indicating that it produces well-reasoned but
less practically relevant tags.

%----------------------------------------------------------------------
\section{Semantic Similarity Analysis}
\label{sec:res-semantic}
%
To complement the token-level and judge-based evaluations, a semantic
similarity analysis compares each model's output to the original ticket
content using four embedding models: OpenAI, E5, SBERT, and BGE.  An
ensemble average is computed across all four embeddings.

\subsection{Problem/Solution Similarity}
%
Table~\ref{tab:semantic-ps} shows the semantic similarity scores for the
Problem/Solution extraction task.

\begin{table}[H]
    \centering
    \caption[Semantic similarity of Problem/Solution output to ticket content]{Semantic similarity of Problem/Solution output to ticket content
             (ensemble average of four embedding models).}
    \label{tab:semantic-ps}
    \begin{tabular}{llccccc}
        \hline
        \textbf{Model} & \textbf{Source} & \textbf{OpenAI} & \textbf{E5} & \textbf{SBERT} & \textbf{BGE} & \textbf{Ensemble} \\
        \hline
        gpt54\_mini  & OpenAI & 0.493 & 0.853 & 0.265 & 0.684 & 0.574 \\
        gpt5.5       & OpenAI & 0.491 & 0.852 & 0.265 & 0.676 & 0.571 \\
        gpt4o\_mini  & OpenAI & 0.449 & 0.845 & 0.260 & 0.674 & 0.557 \\
        gptoss       & Ollama & 0.417 & 0.840 & 0.253 & 0.666 & 0.544 \\
        llama3.1     & Ollama & 0.394 & 0.837 & 0.250 & 0.673 & 0.538 \\
        qwen3\_14b   & Ollama & 0.389 & 0.836 & 0.233 & 0.660 & 0.529 \\
        gemma4       & Ollama & 0.366 & 0.829 & 0.225 & 0.654 & 0.518 \\
        \hline
    \end{tabular}
\end{table}

\subsection{Blind Tag Similarity}
%
Table~\ref{tab:semantic-bt} shows the semantic similarity scores for the
blind tag generation task.

\begin{table}[H]
    \centering
    \caption{Semantic similarity of blind tag output to ticket content.}
    \label{tab:semantic-bt}
    \begin{tabular}{llccccc}
        \hline
        \textbf{Model} & \textbf{Source} & \textbf{OpenAI} & \textbf{E5} & \textbf{SBERT} & \textbf{BGE} & \textbf{Ensemble} \\
        \hline
        gpt5.5       & OpenAI & 0.334 & 0.828 & 0.235 & 0.654 & 0.513 \\
        gpt54\_mini  & OpenAI & 0.311 & 0.825 & 0.226 & 0.654 & 0.504 \\
        gpt4o\_mini  & OpenAI & 0.307 & 0.826 & 0.221 & 0.656 & 0.502 \\
        llama3.1     & Ollama & 0.285 & 0.825 & 0.206 & 0.644 & 0.490 \\
        gptoss       & Ollama & 0.270 & 0.823 & 0.180 & 0.642 & 0.479 \\
        gemma4       & Ollama & 0.231 & 0.825 & 0.130 & 0.636 & 0.456 \\
        qwen3\_14b   & Ollama & 0.216 & 0.823 & 0.119 & 0.633 & 0.448 \\
        \hline
    \end{tabular}
\end{table}

\subsection{Combined Semantic Average}
%
Table~\ref{tab:combined-avg} aggregates the Problem/Solution and Blind Tag ensemble
averages into a single combined score.

\begin{table}[H]
    \centering
    \caption[Combined semantic similarity average]{Combined semantic similarity average (Problem/Solution + Blind Tag).}
    \label{tab:combined-avg}
    \begin{tabular}{llccc}
        \hline
        \textbf{Model} & \textbf{Source} & \textbf{Problem/Solution} & \textbf{Blind Tag} & \textbf{Combined Avg} \\
        \hline
        gpt5.5       & OpenAI & 0.571 & 0.513 & 0.542 \\
        gpt54\_mini  & OpenAI & 0.574 & 0.504 & 0.539 \\
        gpt4o\_mini  & OpenAI & 0.557 & 0.502 & 0.530 \\
        llama3.1     & Ollama & 0.538 & 0.490 & 0.514 \\
        gptoss       & Ollama & 0.544 & 0.479 & 0.511 \\
        qwen3\_14b   & Ollama & 0.529 & 0.448 & 0.489 \\
        gemma4       & Ollama & 0.518 & 0.456 & 0.487 \\
        \hline
    \end{tabular}
\end{table}

The semantic similarity analysis reveals a consistent ranking: proprietary
models lead, followed by \texttt{gptoss} and \texttt{llama3.1} as the
strongest open-source contenders.  Comparing the two tasks, the
Problem/Solution outputs are on average more similar to the ticket content
than the Blind Tag outputs (an average ensemble score of~0.55 across all
models for Problem/Solution, versus~0.49 for Blind Tag).  This gap suggests
that extracting a faithful problem description is easier for the models than
generating appropriate categorical tags, which require mapping free text onto
a fixed taxonomy.

%----------------------------------------------------------------------
\section{Open-Source Model Comparison}
\label{sec:res-oss}
%
Based on the intermediate results from all three evaluation protocols, two
models were selected for the final comparison: \texttt{gpt-oss} and
\texttt{gemma4}.  GPTOSS was chosen because it offers the strongest overall
balance across the protocols: it ranked first in the LLM-as-judge relative
composite, fourth in F1~(0.3562), and fifth in combined semantic
similarity~(0.511).  GEMMA4 was chosen as a contrasting recall-oriented
candidate; although it ranked only fifth in F1~(0.3122) and seventh in semantic
similarity~(0.487), it placed second in the judge composite and achieved the
highest recall among all open-source models~(0.4951).

Among the Ollama-served models, GPTOSS consistently leads the judge rankings
while GEMMA4 excels at recall-oriented tasks.  These two models are carried
forward into the full-corpus run described in Section~\ref{sec:res-fullrun}.

%----------------------------------------------------------------------
\section{Scaling Up: Running the Best Model on the Whole Archive}
\label{sec:res-fullrun}
%
Everything reported so far was measured on the carefully annotated set of
250~benchmark tickets.  That benchmark exists to answer a narrow question:
which model is best?  Once that question is settled, a second and more practical
question follows naturally: what happens when the winning model is let loose on
the company's entire ticket archive, not just the curated sample?  This section
describes that final, large-scale run.

The best-performing configuration from the benchmark was applied to a large
slice of the 2025~data, processing 2{,}742 of the 5{,}730 available
tickets~(roughly~48\,\%).  The purpose of this run is twofold.  It produces a
fully tagged dataset that the company can actually use for downstream analysis,
and it lets us observe how the model behaves on the messier, more varied tickets
that never made it into the clean benchmark.

\subsection{How Fast and How Expensive Was It?}
%
\textit{To be reported once the full run is complete.}  The practical questions
here are how long the model takes per ticket, how many GPU-hours the whole run
consumes, how much memory it needs at peak, and how often the model returns
clean, well-formed JSON on the first try versus how often its output has to be
repaired before it can be used.

\subsection{Which Tags Show Up Most Often?}
%
\textit{To be reported once the full run is complete.}  Counting how often each
of the 126~approved tags appears across the full archive will reveal which kinds
of service problems dominate in practice, and whether the real-world mix of
tickets looks like the balanced benchmark or leans heavily towards a few common
categories.

\subsection{Benchmark vs Full Run: Trend Analysis}
%
\textit{To be reported once the full run is complete.}  Two comparisons are of
interest here.  The first is whether the trend observed in the benchmark across
the three categories, in particular the variety of tags, is proportional to
what is seen on the full run.  The second is whether the semantic similarity
between the model output and the ticket content stays at the same level as on
the benchmark, or changes once the model is applied to the full, unfiltered
archive.

%----------------------------------------------------------------------
\section{Observations}
\label{sec:observations}
%
\subsection{Over-tagging Patterns}
%
The dominant error pattern across all models is over-tagging: models
consistently predict more tags than the human annotator assigned.  Smaller
models~(\texttt{llama3.2}, \texttt{gpt-3.5-turbo}) exhibit over-tagging
rates above~60\,\%, often including loosely related tags that reflect
general domain knowledge rather than ticket-specific evidence.  Larger
models are more conservative and stay closer to the ground-truth tag count.
\texttt{Gemma4} is particularly aggressive: despite ranking 5th in F1, it
achieves the highest recall (0.4951) by assigning many tags, most of which
are imprecise.

\subsection{Under-tagging Patterns}
%
Under-tagging is less common but more pronounced in the LB~(Lagerbelastung)
category.  Warehouse-charge tickets tend to be sparse in narrative text and
heavy in structured codes.  When the narrative signal is weak, models
fail to connect the codes to a meaningful service category and omit tags
that a human recognises from domain context.  \texttt{Qwen3\_14b} and
\texttt{qwen2.5} show the most conservative tagging behaviour with recall
values below~0.26.

\subsection{Category-level Observations}
%
Performance varied markedly across the three ticket categories.  The
Gutschrift~(GS) category, which is also the most frequent, was handled best by
all models, most likely because invoice and credit-note language is well
represented in the pre-training corpora.  The Belastungsanzeige~(BL) category
proved of intermediate difficulty: the charge-notification genre is less common
in open-web training data, and models frequently struggled to distinguish a
requested charge from a confirmed one, which lowered precision.  The
Lagerbelastung~(LB) category was the most challenging of the three.  These
warehouse-charge tickets often consist of technical codes with very little
narrative text, leaving the model little signal to reason from, and they
consequently exhibited the highest under-tagging rate as models failed to
infer latent service actions from article numbers alone.

\subsection{Tag Normalisation and Surface-Form Errors}
\label{sec:results-language}
%
Because the F1 evaluation compares tag strings, it is sensitive to superficial
differences in how a tag is written even when the meaning is identical.  Four
distinct sources of such surface-form mismatch were observed, and each had to
be normalised before a fair comparison was possible.

The first is language.  Although all models are instructed to produce English
output, the smaller models occasionally respond in German when the input is
entirely in German.  German compound words such as \textit{Lagerbelastung} are
then reproduced verbatim instead of being mapped to the approved English tag,
which lowers precision.  Given the benchmark language
distribution~(82.8\,\% German, 15.7\,\% English, and a small remainder spread
across a few other languages such as Russian, Spanish, and Estonian), the large
German majority makes this code-switching problem common.

The second is grammatical number: a model might emit \texttt{missing\_part} in
one ticket and \texttt{missing\_parts} in another, so singular and plural forms
of the same concept were folded together.

The third is the choice of separator.  The same tag was variously written with
a hyphen, a space, or an underscore between words; all separators were
normalised to a single canonical form using the underscore, so that
\texttt{spare\_part}, \texttt{spare-part}, and \texttt{spare part} are treated
as one tag.

The fourth, and most subtle, is synonymy: two different tags sometimes express
the same underlying meaning through synonyms, and these were consolidated onto
a single canonical label so that semantically identical tags were no longer
counted as mismatches.

After applying all four normalisation steps, together with lower-casing, the
predicted tags can be compared against the ground truth on a meaning-preserving
basis rather than being penalised for purely cosmetic differences.
