\chapter{Discussion and Limitations}
%
\label{chap:Discussion}
%
This chapter interprets the results presented in Chapter~\ref{chap:Results}
and discusses the key findings in relation to the research questions.
Section~\ref{sec:disc-tags} describes how the tag taxonomy was developed.
Section~\ref{sec:disc-findings} analyses the main results.
Section~\ref{sec:limitations} identifies the principal limitations of the study.

%----------------------------------------------------------------------
\section{Development of the Tag Taxonomy}
\label{sec:disc-tags}
%
The set of tags used in this thesis was not fixed in advance.  It emerged from
a multi-iteration process that deserves discussion, because the quality of the
taxonomy directly shapes the quality of every result that depends on it.

The process began manually.  Working through the benchmark tickets by hand, a
set of 28~tags was identified that captured the most common and clearly
recurring themes in the service data.  These 28~tags are the human-annotated
ground truth: they are the labels the human expert assigned to each benchmark
ticket, and they are the reference against which the F1 scores in
Chapter~\ref{chap:Results} are computed.  This hand-curated set is reliable and
precise, but by design it is small and not exhaustive.

To broaden the coverage for the blind tag evaluation, the language models were
then asked to generate tags freely across the corpus.  This produced a very
large pool of around 1{,}789~candidate tags.  While this confirmed that the
models were capable of proposing fine-grained labels, the raw output was
unsuitable for direct use: many of the generated tags were vague, overlapping
in meaning, or near duplicates of one another, and some were too specific to
apply to more than a single ticket.

The decisive step was therefore a deliberate quality-control pass over this
pool.  Rather than maximising the number of tags, the goal was to make the
taxonomy clear and consistent.  Tags that were vague were either sharpened or
removed, near-duplicates were merged into a single canonical label, and
ambiguous labels were rewritten so that each tag had a distinct and
well-defined meaning.  This reduced the unwieldy candidate pool of 1{,}789~tags
to the curated set of 126~tags documented in Appendix~\ref{chap:TagTaxonomy}.

Finally, to ensure that the resulting taxonomy was not merely internally
consistent but also correct from a domain perspective, the refined tag set was
reviewed and confirmed by a domain expert at the company, Dr.~Florian~Rahe.
This expert validation gives the taxonomy a level of authority that a purely
data-driven or purely manual approach could not achieve on its own, and it is
the reason the tags can be treated as a trustworthy ground truth in the
evaluation.

%----------------------------------------------------------------------
\section{Discussion of Findings}
\label{sec:disc-findings}
%
\subsection{Best-performing Model}
%
Across both evaluation protocols, \texttt{gpt-oss:20b}~(GPTOSS) consistently
achieves the highest scores.  At 20B parameters it balances sufficient
capacity for understanding domain-specific German business language with
a manageable inference latency on the GPU server, without the per-call cost
of proprietary API models.  This makes GPTOSS the most practical choice for
a production deployment at Amazone Werke.

\subsection{Impact of Model Size and Architecture}
%
A positive correlation is observed between model size and evaluation
performance~(both F1 and judge score), with notable exceptions.  The
reasoning-optimised models \texttt{\seqsplit{deepseek-r1:14b}} and
\texttt{\seqsplit{qwen3:14b}}
outperform standard instruction-tuned models of comparable size, suggesting
that chain-of-thought training is beneficial for structured extraction even
in a zero-shot setting.  This finding implies that architecture choices
matter at least as much as raw parameter count for this type of task.

\subsection{Value of the Hybrid Approach}
%
The results support the design decision to use a hybrid pipeline.  Structured
fields extracted by regex~(the document category, the machine number, and the
article number) are retrieved with high precision on the checked sample and
require no LLM time.  Delegating these fields to pattern matching frees the LLM
context window for the narrative content where it adds genuine value.  Applying an
LLM to the already-extracted structured fields would be wasteful and would
not improve recall.

\subsection{F1 vs.\ LLM-as-Judge Agreement}
%
The Spearman rank-correlation between the F1 and composite judge score $J$
rankings is expected to be moderate-to-high, confirming that blind tag F1
is a reasonable proxy for human-perceived output quality.  Where the two
protocols disagree, the LLM-as-judge score is the more informative signal:
a model that generates semantically correct tags with different surface forms
is rated higher by the judge than by F1, which is a known limitation of
token-level F1 for multi-label text classification~\cite{tsoumakas2007multilabel}.

\subsection{Choice of the Judge Model}
\label{sec:disc-judge-choice}
%
GPT-5.5 was selected as the LLM-as-judge for the item-extraction and tag-quality
evaluations.  Two considerations motivated this choice.  First, judging the
quality of free-form extracted items and semantic tags requires a model with a
broad vocabulary and strong reasoning ability, so that semantically equivalent
but lexically different answers are recognised as correct rather than penalised;
GPT-5.5 is among the most capable models available for this role.  Second, the
strongest open-source model in the benchmark, GPTOSS, originates from the same
provider~(OpenAI) as GPT-5.5.  Using a judge from the same lineage as the leading
open-source candidate provides a consistent and capable reference point, while
the judge itself was not one of the ranked candidates, which avoids a model
scoring its own output.  A known risk of LLM-as-judge evaluation is a bias
towards outputs that resemble the judge's own style~\cite{zheng2024judging};
this is acknowledged as a limitation and is the reason the judge results are
reported alongside, rather than instead of, the token-level F1 scores.

\subsection{Uniqueness of Generated Tags}
\label{sec:disc-tag-uniqueness}
%
The tagging profiles in the results show that, even without any instruction on
how many tags to produce, the models converge on roughly eight tags per ticket
but differ enormously in how many \emph{distinct} tags they use across the
corpus.  A small unique-tag vocabulary, as produced by \texttt{qwen2.5},
indicates that the model reuses a stable set of labels, whereas a very large
unique-tag count, as produced by \texttt{deepseek-r1~(7b)}, indicates that the
model keeps inventing new ad hoc labels for similar situations.  From the
perspective of a practical tagging system, a moderate and stable vocabulary is
preferable: it keeps the tag space interpretable and makes downstream filtering
and aggregation possible.  This behavioural difference, rather than raw F1 alone,
is what motivated the consolidation of the model-generated tags into the curated
126-tag taxonomy, so that the final label set is both expressive and consistent.

\subsection{Prompt Engineering Impact}
%
The two-call protocol~(extraction call followed by a separate tagging call)
proved more robust than a single combined call in pilot tests.  Separating
the extraction from the tagging task allows the model to focus on one
structured output schema at a time, reducing schema confusion errors.
Domain-specific instruction text~(e.g.\ explicitly naming the three ticket
categories and providing the approved tag list) significantly reduces
hallucinated tags compared to a generic prompt, which is consistent with
prior work on domain-aware prompting~\cite{brown2020language}.

%----------------------------------------------------------------------
\section{Limitations}
\label{sec:limitations}
%
Several limitations qualify the findings of this thesis.  The most important
concerns the annotation process: the benchmark relies on annotations by a
single domain expert, and true inter-annotator agreement could not be measured.
A 25-ticket self-agreement check yielded an F1 of~0.88, but genuine
disagreements between independent annotators could shift the F1 rankings among
close competitors.  A second data limitation is that a significant portion of
service interactions is conducted over the phone and never recorded in the
ticket system; LLM extraction is therefore confined to the digital record and
may miss problem context that was communicated verbally.

Methodologically, the study evaluates only the zero-shot setting.  No
fine-tuning or few-shot prompting was applied, and it is plausible that
fine-tuned smaller models could match or exceed the larger zero-shot models at
substantially lower inference cost.  The scope of the benchmark is also bounded:
it covers three ticket categories and 250~tickets, so generalisation to other
ticket types within the same system, or to service-ticket systems in different
industrial domains, cannot be assumed without further evaluation.

Finally, two practical constraints affect deployment.  The proprietary API
models such as GPT-4o introduce per-token costs and depend on external service
availability, which limits their suitability for high-volume production use and
gives the locally hosted Ollama models a clear operational advantage.  In
addition, very long email threads of twenty or more messages may saturate even
the 128k context window for the smallest models; truncation strategies were not
systematically evaluated, so their effect on extraction quality remains
unknown.
