\chapter{Introduction}
%
This document template can be used as a model for academic theses. It is based on 
guidelines created by Prof.\,Dr~Stephan Freichel as Chairman of the Examination 
Board for the B.\,Sc. Logistics and M.\,Sc. Supply Chain and Operations 
Management degree programs at the Faculty of Automotive Systems and Production.

The text in this template describes general formal requirements, in particular 
for the table of contents, the insertion of references and the compilation of a 
bibliography.

The template can be used interdisciplinary as a sample file for theses at TH 
Köln. However, when using it, you must make sure that it complies with the 
conventions of your degree program.

The content of the document was revised and modified by Maria-Anna Worth 
(retired) and Susanne Neuzerling (Department of Planning and Controlling) to make 
it suitable for use across all subjects. A further revision and update was 
carried out by Andreas Bissels (Writing Centre). Ms. Katharina Bata created the 
document template in LaTeX. This template was last revised in 2023 by André 
Ulrich and Jan Salmen and supplemented with various notes, specifically on design 
with \LaTeX{} (see \cref{chap:Textsatz}). The document was translated by Lynne 
Lörler (Language Learning Center).

To prepare for your final thesis, we recommend that you take advantage of the 
services offered by the Writing Center of the Center for Soft Skills Development%
\footnote{\href{https://www.th-koeln.de/schreibzentrum}{https://www.th-koeln.de/schreibzentrum}}%
; these include both counselling%
\footnote{\href{https://www.th-koeln.de/schreibberatung}{https://www.th-koeln.de/schreibberatung}}
and courses%
\footnote{\href{https://www.th-koeln.de/schreibzentrum/seminare}{https://www.th-koeln.de/schreibzentrum/seminare}}
on academic writing. The Writing Center is 
your point of contact at TH Köln for questions relating to academic writing.
Additionally, the Language Learning%
\footnote{\href{https://www.th-koeln.de/sprachlernzentrum}{https://www.th-koeln.de/sprachlernzentrum}}
Center offers counselling%
\footnote{\href{https://www.th-koeln.de/internationales/schreibberatung-englisch-fuer-studierende_84724.php}{https://www.th-koeln.de/internationales/schreibberatung-englisch-fuer-studierende\_84724.php}}
on writing in English.

Please save this document template twice on your computer: once so that you can 
continue to access the content presented here, and a second time so that you can 
overwrite it with your own thesis.

\emph{Please note: This template does not supersede the specific requirements of the respective examination boards. If there are special formal requirements in your subject, these apply.}%\enlargethispage{\baselineskip}
%
\begin{flushright}
Cologne, November 2024
\end{flushright}