\chapter{Related Work}
%
\label{chap:RelatedWork}
%
This chapter surveys the research areas most directly relevant to this thesis.
It begins with a broad view of how large language models are applied to
real-world analysis tasks~(Section~\ref{sec:rw-llm-apps}), narrows to their
use for converting unstructured into structured data~(Section~\ref{sec:rw-struct}),
and then covers the four areas that define the specific contribution of this
work: (1)~the use of LLMs for information extraction~(Section~\ref{sec:rw-llm-ie}),
(2)~automated service ticket analysis~(Section~\ref{sec:rw-tickets}),
(3)~multi-label text classification~(Section~\ref{sec:rw-multilabel}), and
(4)~multilingual NLP~(Section~\ref{sec:rw-multilingual}).  It closes with the
LLM-as-judge evaluation paradigm~(Section~\ref{sec:rw-judge}) and a summary of
the research gap~(Section~\ref{sec:rw-gap}).

%----------------------------------------------------------------------
\section{Applications of Large Language Models for Analysis Tasks}
\label{sec:rw-llm-apps}
%
Large language models have rapidly moved beyond text generation into a wide
range of analytical and operational tasks.  One particularly active area is
\emph{root cause analysis}~(RCA), where the goal is to diagnose the underlying
reason for a system failure or incident from heterogeneous textual evidence.
This is structurally similar to diagnosing the root problem of a service
ticket.  \citeauthor{chen2024automatic}~\cite{chen2024automatic} applied LLMs
to cloud incident reports with their \textsc{RCACopilot} system, which uses
predefined handlers to automatically collect multi-modal diagnostic data and
an LLM to predict a root cause category, avoiding the cost of repeated
fine-tuning.  Building on this, \citeauthor{roy2024exploring}~%
\cite{roy2024exploring} explored \emph{agentic} LLM systems on an
out-of-distribution dataset of production incidents collected at Microsoft.
Their central finding is instructive for this thesis: a ReAct agent equipped
with retrieval tools performed competitively with strong retrieval and
reasoning baselines \emph{but with substantially higher factual accuracy}, and
they observed that introducing explicit chain-of-thought reasoning yielded the
lowest hallucination rates.  Notably, they also report that adding more context
(historical incident discussions) did \emph{not} yield significant performance
improvements, a cautionary result for context-stuffing approaches.
\citeauthor{pei2025flowofaction}~\cite{pei2025flowofaction} extended the agent
paradigm with standard-operating-procedure guidance to constrain the diagnostic
search in a multi-agent setting.

These works are relevant to this thesis because they share the same fundamental
challenge: extracting an actionable diagnosis from noisy, semi-structured
operational text.  The service tickets analysed in this thesis differ in domain
(agricultural machinery rather than cloud infrastructure) and in language
(predominantly German rather than English), but the core task of distilling a
coherent problem and solution from a messy thread is closely analogous to RCA.
The RCA findings on the value of explicit reasoning and the diminishing returns
of additional context directly informed the prompt design choices made in
Chapter~\ref{chap:Methodology}.

\subsection{Reasoning Failures and Domain Adaptation in RCA}
\label{subsec:rw-rca-failures}
%
A more recent line of work moves beyond demonstrating that LLMs \emph{can}
perform RCA and instead asks \emph{why} they fail, which is directly relevant
to interpreting the error patterns observed in this thesis.
\citeauthor{kim2026why}~\cite{kim2026why} run a controlled study of LLM-based
cloud RCA and distil a taxonomy of twelve recurring pitfall types.  Their
central conclusion is sobering for prompt-only approaches: prompt engineering
alone does not resolve the dominant reasoning failures, and improvements in how
agents communicate with one another help more than refinements to a single
prompt.  In a similar spirit, \citeauthor{riddell2026stalled}~%
\cite{riddell2026stalled} build a taxonomy of reasoning failures for
cloud-based RCA, compare ReAct against Plan-and-Execute workflows across six
models, and show that the type of reasoning failure is itself predictive of
whether the final answer is correct.  Together these studies suggest that the
over-tagging and under-tagging patterns reported in Chapter~\ref{chap:Results}
are better understood as reasoning failures than as simple labelling noise.

Domain adaptation emerges as the most reliable remedy.
\citeauthor{sana2025reasoning}~\cite{sana2025reasoning} benchmark RCA in 5G
wireless networks and find that open-source reasoning models struggle without
domain-specific adaptation; a two-stage fine-tuning approach combining
supervised learning with reinforcement learning markedly improves both accuracy
and the quality of the reasoning.  \citeauthor{tan2026llmrca}~%
\cite{tan2026llmrca} take a complementary, retrieval-augmented route with their
multilevel \textsc{LLMRCA} framework, which fuses multimodal observability data
into a heterogeneous causal graph and outperforms eight baselines on
response-quality problems.  The recurring lesson, that generic prompting is not
enough and that domain grounding is decisive, reinforces the design decision in
this thesis to embed Amazone-specific vocabulary and an expert-validated tag
taxonomy directly into the pipeline rather than relying on a general-purpose
prompt.

\subsection{From Correlation to Causation: Prompting for Causal Reasoning}
\label{subsec:rw-causal}
%
Underlying all RCA is the more fundamental question of whether LLMs can reason
causally at all rather than merely echoing surface correlations.  Two studies
are particularly relevant.  \citeauthor{sgouritsa2024prompting}~%
\cite{sgouritsa2024prompting} introduce \textsc{PC-SubQ}, a prompting strategy
that decomposes causal inference into a fixed sequence of sub-questions aligned
with a classical causal-discovery algorithm; across five LLMs it improves on
baseline prompting and, importantly, stays robust when variable names are
changed or the question is paraphrased.  \citeauthor{liu2025eliciting}~%
\cite{liu2025eliciting} approach the problem from a different angle and find
that code-style prompts and code-trained LLMs reason about cause and effect
more reliably than text-only prompting, and that fine-tuning on conditional
(if-then) statements improves causal reasoning in both code and natural
language.  These findings matter for the present work because assigning a
service tag is implicitly a causal judgement: the model must infer, from a
description of symptoms, the underlying fault that caused them.  The evidence
that structured, decomposed prompts improve causal inference supports the
explicit two-call protocol adopted here, in which problem and solution are
extracted before tags are assigned.

%----------------------------------------------------------------------
\section{Structuring Unstructured Data with LLMs}
\label{sec:rw-struct}
%
A second broad line of research focuses on converting unstructured or
semi-structured data into structured representations.  Recent work on document
understanding shows a consistent hybrid pattern: optical character recognition,
parsing, or rule-based steps handle the first pass, and LLMs are then used for
semantic extraction, normalisation, or schema-aware cleanup.
\citeauthor{stuhlmiller2025scalable}~\cite{stuhlmiller2025scalable} report that
an LLM pipeline can process semi-structured and unstructured electronic health
records to improve classification, extraction, and standardisation.  This
mirrors the hybrid design adopted in this thesis, where reliable regular
expressions capture structured fields and the LLM reasons over the unstructured
narrative.

Benchmark-oriented studies, however, caution that extraction remains fragile
when documents become messy or schemas become broad.
\citeauthor{deng2025unstructured}~\cite{deng2025unstructured} present a
comprehensive benchmark for unstructured data analysis with LLMs, and
\citeauthor{ferguson2026extractbench}~\cite{ferguson2026extractbench} introduce
an evaluation methodology specifically for complex structured extraction, both
finding substantial performance gaps as document complexity grows.  Work on
tabular comprehension is also directly relevant:
\citeauthor{sui2024table}~\cite{sui2024table} designed the \textsc{SUC}
benchmark of seven structural-understanding tasks~(e.g.\ cell lookup, row
retrieval, size detection) and evaluated GPT-3.5 and GPT-4 on it.  Their
results show that LLMs only partially understand structured table data and that
performance depends heavily on input design: serialising tables as HTML
outperformed natural-language-with-separators by roughly~6.8\%, and their
proposed \emph{self-augmented prompting}~(asking the model to first describe the
table's structure, then answer) raised overall accuracy to about~65\% across
the seven tasks.  \citeauthor{he2024thinking}~\cite{he2024thinking} similarly
demonstrate that supplying explicit structure~(in the form of knowledge graphs)
improves reasoning.  These findings carry a clear lesson for this thesis: the
\emph{format} in which ticket content is presented to the model matters as much
as the model itself, which is why the preprocessing pipeline converts all
sources into a consistent Markdown representation~(Chapter~\ref{chap:Methodology}).

%----------------------------------------------------------------------
\section{Large Language Models for Information Extraction}
\label{sec:rw-llm-ie}
%
Since the introduction of the Transformer architecture~\cite{vaswani2017attention},
pre-trained language models have become the dominant paradigm for NLP tasks.
\textsc{BERT}~\cite{devlin2019bert} demonstrated that fine-tuning a single
pre-trained model achieves strong results across diverse extraction tasks.
More recently, instruction-tuned decoder-only models such as the
GPT~\cite{brown2020language} and Llama~\cite{touvron2023llama} families have shown
that well-crafted zero-shot prompts can match or surpass fine-tuned baselines on
many information extraction benchmarks.

\citeauthor{wei2022finetuned}~\cite{wei2022finetuned} showed that instruction
fine-tuning substantially improves zero-shot performance, while
\citeauthor{kojima2022large}~\cite{kojima2022large} demonstrated that chain-of-thought
prompting further unlocks reasoning in larger models.  Both findings motivate the
structured, step-by-step prompt design used in this thesis
(see Section~\ref{sec:prompt-design}).

Domain-aware prompts consistently outperform generic prompts.
\citeauthor{keloth2024advancing}~\cite{keloth2024advancing} showed for biomedical
named entity recognition that instruction-tuned LLMs adapted to domain vocabulary
substantially improve extraction quality, and similar results have been reported
in legal document analysis.
Accordingly, the prompts in this thesis embed Amazone-specific domain vocabulary and
define a fixed output schema.

%----------------------------------------------------------------------
\section{Automated Service Ticket Analysis}
\label{sec:rw-tickets}
%
Automated analysis of IT helpdesk tickets has been studied since at least the
early 2000s.  Classical approaches relied on keyword matching and TF-IDF
vectorisation combined with support vector machines for ticket categorisation
\cite{yang1999reevaluation}.  More recent work applies deep learning: recurrent
networks \cite{hochreiter1997long} and BERT-based classifiers are now common for
ticket routing and priority prediction.

However, the bulk of this literature targets English-language IT helpdesk data
from publicly available datasets~(e.g.~Incident Ticket from ServiceNow).
A dedicated line of work focuses narrowly on service tickets and email triage,
typically casting the task as classification, prioritisation, or escalation.
\citeauthor{tolciu2021analysis}~\cite{tolciu2021analysis} analyse patterns and
similarities in service tickets using classical NLP techniques such as
bag-of-words and TF-IDF.  \citeauthor{liu2023ticketbert}~\cite{liu2023ticketbert}
introduces \textsc{Ticket-BERT}, fine-tuning a BERT-style model to label incident
management tickets, and \citeauthor{boonklay2025comparative}~%
\cite{boonklay2025comparative} provide a comparative study of machine learning
and deep learning algorithms for customer support ticket classification across
public and enterprise datasets.  These papers show that ticket automation is
feasible and often accurate, but they mostly optimise routing or label
assignment rather than extracting and synthesising actionable insights from the
combination of emails, PDFs, and attachments.

Industrial service tickets for physical machinery, which are multilingual,
attachment-heavy, and span years of email threads, are comparatively
understudied.  The closest related work by
\citeauthor{hassan2023llm}~\cite{hassan2023llm}
applies GPT-4 to manufacturing quality reports but does not address the
extraction of structured entities or multi-label categorisation.

This yields a clear research gap: there is substantial prior work on either
rule-based extraction from structured document fields or ML/LLM classification
of support tickets, but less work that explicitly combines reliable rule-based
capture for the structured parts with LLM reasoning over the unstructured
narrative to produce coherent operational insights.  This thesis contributes to
filling this gap by constructing a benchmark for agricultural machinery service
tickets and evaluating multiple LLMs on it.

%----------------------------------------------------------------------
\section{Multi-label Text Classification}
\label{sec:rw-multilabel}
%
Service tickets require multi-label classification because a single ticket may
simultaneously involve a warranty claim, a spare part request, and an electronic
fault.  Classical multi-label learning surveys~\cite{zhang2014review} distinguish
problem-transformation methods~(e.g.\ binary relevance, label powerset) from
algorithm-adaptation methods.

Transformer models reformulate multi-label classification as sequence generation,
outputting a comma-separated list of labels or a JSON array.  This generative
framing is increasingly applied to industrial data: \citeauthor{llm4industrial2024}~%
\cite{llm4industrial2024} use LLMs for multi-label classification of industrial
requirements and observe that parameter count alone does not determine
performance, with smaller well-tuned models competing with much larger ones.
The blind-tag evaluation protocol used in this thesis follows the same
generative multi-label paradigm, which is well suited to settings where the
label set is large and labelled training data is scarce, precisely the setting
of this thesis.

An important distinction is between \emph{accuracy-oriented} evaluation~(strict
match against a fixed ground truth) and \emph{usefulness-oriented} evaluation.
The latter is more appropriate when multiple valid labellings exist, and is
captured here through the LLM-as-judge protocol described in
Section~\ref{sec:res-judge}.

%----------------------------------------------------------------------
\section{Multilingual NLP and Code-switching}
\label{sec:rw-multilingual}
%
Service tickets in this corpus are predominantly in German but frequently contain
English phrases, Latin abbreviations, and brand-specific terminology.
This phenomenon, known as \emph{code-switching}, poses challenges for models
trained predominantly on English data.

Multilingual models such as mBERT~\cite{devlin2019bert} and
XLM-R~\cite{conneau2020unsupervised} handle code-switching with varying success
depending on the language pair and domain.  Among the LLMs evaluated in this
thesis, instruction-tuned models with multilingual training data~(the Gemma
and Qwen families) generally outperform English-centric models on German
passages.

%----------------------------------------------------------------------
\section{LLM Evaluation and the LLM-as-Judge Framework}
\label{sec:rw-judge}
%
Standard NLP metrics such as F1-score and exact match are insufficient for
evaluating free-form generation where multiple valid outputs exist.
\citeauthor{zheng2024judging}~\cite{zheng2024judging} introduced the
LLM-as-judge paradigm, in which a capable model~(e.g.~GPT-4) evaluates the
quality of outputs from other models using structured rubrics.  This approach
has been adopted for summarisation, dialogue, and, more recently, information
extraction evaluation.

This thesis applies LLM-as-judge to rate tag usefulness and reasoning coherence,
complementing the F1-based evaluation against human annotations.

%----------------------------------------------------------------------
\section{Summary and Research Gap}
\label{sec:rw-gap}
%
Table~\ref{tab:rw-gap} summarises the positioning of this thesis relative to
prior work.

\begin{table}[H]
    \centering
    \caption{Positioning of this thesis in the related-work landscape.}
    \label{tab:rw-gap}
    \begin{tabular}{lccccc}
        \hline
        \textbf{Work} & \textbf{LLM} & \textbf{Industrial} & \textbf{Multilingual}
                      & \textbf{Multi-label} & \textbf{Benchmark} \\
        \hline
        Yang~et~al.~\cite{yang1999reevaluation}   & No  & No  & No  & Yes & Yes \\
        Hassan~et~al.~\cite{hassan2023llm}         & Yes & Yes & No  & No  & No  \\
        Industrial ML~\cite{llm4industrial2024}        & Yes & Yes & No  & Yes & No  \\
        \textbf{This thesis}                        & Yes & Yes & Yes & Yes & Yes \\
        \hline
    \end{tabular}
\end{table}

The combination of an industrial, multilingual setting with multi-label
classification and a human-annotated benchmark represents a contribution not
addressed by any single prior work.
