\chapter{Introduction}
%
\label{chap:Introduction}
%
\section{Motivation}
%
Modern industrial companies accumulate thousands of service tickets every year.
A single ticket at an agricultural machinery manufacturer such as Amazone~Werke
can span dozens of email exchanges, multi-page PDF invoices, spreadsheet attachments,
and scanned delivery notes, all in a mix of German, English, and occasionally French
or Dutch. Extracting actionable information from this heterogeneous corpus manually
is slow, inconsistent, and impossible to scale.

Advances in Large Language Models~(LLMs) offer a promising alternative. Models
such as Llama~3, Gemma~3, and GPT-4o have demonstrated strong zero-shot
capabilities in information extraction and text classification, even in specialised,
domain-specific settings. However, applying these models to real-world industrial
service data, with its multilingual content, verbose email threads, and
domain-specific abbreviations, requires a principled experimental evaluation that
is largely absent from the academic literature.

This thesis addresses that gap. It designs and evaluates an end-to-end pipeline that
(1)~converts raw ticket files into a unified textual representation, (2)~uses LLMs
to extract structured insights~(problem statement, solution, entity identifiers, and
categorical tags), and~(3)~rigorously benchmarks fifteen model configurations
(ten open-source models served locally and five proprietary API models) against
human-annotated ground truth.

%----------------------------------------------------------------------
\section{Problem Statement}
\label{sec:problem}
%
Service tickets at Amazone~Werke exhibit several properties that together make
their automated analysis challenging.  The most fundamental difficulty is that
the content is unstructured and heterogeneous: important details are spread
across email bodies~(MSG), PDF documents such as credit notes and inspection
reports, Excel spreadsheets, and scanned images, and no consistent schema ties
these sources together.  Compounding this, the text is frequently
multilingual.  A single ticket may contain paragraphs in German and English
interleaved with technical abbreviations drawn from both languages, so standard
NLP tools optimised for one language fail to process the code-switched content
reliably.

A further complication is that the information itself is of two distinct kinds.
Document categories~(Gutschrift, Belastungsanzeige), article codes, and machine
serial numbers follow predictable patterns that rule-based extraction handles
well, whereas the underlying problem, its root cause, and the eventual
resolution require contextual reasoning that rules alone cannot provide.  Any
effective system must therefore combine both extraction strategies rather than
relying on a single technique.

Two practical constraints frame the task.  First, the corpus comprises tens of
thousands of tickets, which makes manual analysis infeasible and demands a
system that operates with minimal human intervention.  Second, the written
record is often incomplete: many ticket resolutions are agreed upon in phone
calls or meetings and never fully documented, so a model must reason under
uncertainty without fabricating solutions that the source text does not
support.  The central challenge of this thesis is to extract useful insights
from this material without losing the context that gives those insights
meaning.

%----------------------------------------------------------------------
\section{Research Questions}
\label{sec:rq}
%
The thesis is structured around one main research question and two
sub-questions that refine it.

\begin{description}
    \item[\textbf{RQ\,1 (Main):}] \emph{Can Large Language Models reliably extract
          actionable insights from industrial service tickets?}

    \item[\textbf{RQ\,1.1:}] \emph{Which ticket information can be captured reliably
          with rule-based extraction, and which parts need LLMs?}

    \item[\textbf{RQ\,1.2:}] \emph{Which model produces the most useful and coherent
          insights on the benchmark tickets?}
\end{description}

%----------------------------------------------------------------------
\section{Scope and Delimitations}
\label{sec:scope}
%
This thesis focuses on the \emph{extraction} of structured information from existing
ticket files; it does not cover automated ticket routing, priority prediction, or
resolution recommendation.  The evaluation is conducted on a benchmark of
250~tickets drawn from the Amazone~Werke service corpus; generalisation to other
industrial domains is discussed qualitatively in Chapter~\ref{chap:Conclusion}.

Models are evaluated in a zero-shot configuration using a fixed, domain-aware prompt.
Fine-tuning and retrieval-augmented generation~(RAG) are left as directions for
future work.

%----------------------------------------------------------------------
\section{Thesis Structure}
\label{sec:structure}
%
The remainder of this thesis is organised as follows.
Chapter~\ref{chap:RelatedWork} reviews related work on LLM-based information
extraction, service ticket analysis, and multi-label text classification.
Chapter~\ref{chap:Methodology} describes the overall methodology, including the
preprocessing pipeline, benchmark construction, and the experimental setup.
Chapter~\ref{chap:Results} presents the quantitative results and qualitative
observations.
Chapter~\ref{chap:Discussion} interprets the findings and discusses the
limitations of the approach.
Chapter~\ref{chap:Conclusion} summarises the findings, answers the research
questions, and outlines directions for future work.
